Multiphyscis modeling for the Fluoride-salt-cooled High-temperature Reactor (FHR)

### 2018-07-22 paper/main.tex
Updates from ShareLaTeX

\begin{abstract}
Fluoride-salt-cooled High-temperature Reactor(FHR) is one of the generation IV reactor designs that promises great safety and market competitiveness because of its unique combination of TRISO fuel elements, liquid salt coolant and graphite reflectors.
 The computational ability to accurately predict its dynamic behavior during postulated transients is important for FHR technology development and licensing. This paper describes a fully coupled neutronics and thermal hydraulics modeling methodology based on multi-group neutron diffusion and multi-scale porous media CFD.

 To demonstrate the application and credibility of the method, an example model is built in COMSOL multiphysics for the TMSR design and results for eigenvalue, steady state and transient studies are extensively verify against a reference code , which couples Monte Carlo neutron transport/CFD code and have shown satisfactory agreement.
\end{abstract}

@@

As illustrated in figure \ref{fig:triso}, a commercial scale reactor core would contain tens of thousands fuel pebbles and within each thousands of TRISO particles that encloses the fuel material. This peculiar geometry poses unique challenges to both neutronics and thermal-hydraulics modeling. Reactor-agnostic Monte Carlo methods can model the particles and pebbles explicitly but require a large number of source neutrons to achieve desired statistical accuracy, especially for transient analysis. Deterministic codes, in the contrary, model the fuel region as homogenized medium with dedicated cross-sections generated from other codes, among which diffusion equations compute two-dimensional or three-dimensional neutron behaviour with reasonable computational cost and accuracy, although limitations can be found in vicinity of control rods, burnable poisons, and other abrupt discontinuities.

@@

The question of how to model the complex geometry of the flow passage in a pebble bed has been tackled in chemistry engineering\cite{Dixon2001}\cite{Miroliaei2011} for fixed bed reactors in chemical processes and in nuclear engineering for analyzing advanced pebble bed reactor designs.  One-dimensional heat conduction simplification is often used in LWR fuel rods and can be used in spherical fuel elements for FHRs for quick scope analysis, but is inadequate for detailed investigation because it ignores temperature gradients in the axial and the radial direction, which can be significant in a FHR core. Two main approaches for full core CFD modeling are realistic modeling and porous media modeling. Realistic modeling has been used to model flow passage directly in a segment of pebbles, from several to several hundreds\cite{Lee2007}\cite{Hassana}\cite{Shams2013}, but quickly becomes unfeasible for a large number of pebbles with the current computation resources. In addition, although the goal of realistic modeling is to model the flow passage realistically, it is not free of assumptions\cite{Lee2007}, e.g. pebble packing structure and contact point treatment. The porous media approach, on the other hand, computes only the average flow conditions with ten time less mesh grids. Comparisons between the direct CFD and porous media methods\cite{Wu2010} suggest that porous media methods provide similar averaged pressure and temperature gradient across the pebble bed to the realistic approach, given the appropriate Darcy-Forchheimer drag model is used. Porous media approach has been used in standard HTR modeling tools Thermix\cite{Struth1985} and Tinte\cite{Gerwin1987}, as well as in a thermal-hydraulics model for FHR\cite{Scarlat2012}.
Appropriate safety margin should be applied in the design given the fact that porous media models do not capture local vertices, flow separation, variations in flow induced heat transfer around the pebble surface and the resulting local hot spots.

Previously, researchers have developed separate codes in thermal-hydraulics\cite{Scarlat2012} and neutronics\cite{Cisneros2013} for FHRs. These codes can be coupled for multiphysics computation using the operator-split method: solving one equation system at a time and transfer data between the codes after each iteration until the results converge. However, the operator split method requires small time steps to ensure convergence due to nonlinear inconsistency\cite{Ragusa2009} and could require a substantial amount of computational resource. This work explores methodology for developing fully coupled neutronics and thermal-hydraulics models for pebble bed FHRs.

The rest of the paper is organized as following: The methodology, including governing equations, generation of cross-sections and implementation in COMSOL is discussed in section \ref{sec:methodology}. A case study based on TMSR design is then presented in section \ref{sec:res}, as well as verification procedures against a reference model. Future needs are identified in the conclusion.

@@

\subsection{Neutronics}
Multi-group diffusion equation provides a good balance between CPU cost and accuracy, and is used in the multiphyscis model for neutronics computation. The governing equations of multi-group neutron diffusion with delayed neutrons are written as following:
  \begin{equation}
    \frac{1}{v_g}\frac{\partial \phi_g}{\partial t} =& \nabla D_g \nabla \phi_g - \Sigma_{t,g}\phi_g + \sum_{g'=1}^{G} \Sigma_{s, g'g}\phi_{g'} + \\\nonumber & \ (1-\beta)\chi_{p,g}\sum_{g'=1}^G(\nu \Sigma_f)_{g'} \phi_{g'} + \chi_{d,g}\sum_{i=1}^D \lambda_iC_i
    \label{eq:diffusion}
  \end{equation}

@@

For criticality calculations, the effective multiplication factor \keff is inserted into the neutron balance equation to represent the variation of neutron population from one generation to the following one, resulting in:
  \begin{align}
    0 =& \nabla D_g \nabla \phi_g - \Sigma_{a,g}\phi_g - \sum_{\substack{1\\g'\neq g}}^G \Sigma_{s, gg'}\phi_{g} + \\\nonumber &\sum_{\substack{1\\g'\neq g}} \Sigma_{s, g'g}\phi_{g'} + \frac{\chi_{t,g}}{k_{\mathrm{eff}}}\sum_{g'=1}^G(\nu \Sigma_f)_{g'} \phi_{g'}
  \end{align}

A vacuum boundary condition is applied at the outer surfaces. And a symmetry boundary condition is used at the core center-line in 2-D axi-symmetrical models.

\subsubsection{Cross-section generation}
The group-averaged and component-wise neutronic constants are generated with the Monte Carlo neutron transport code SERPENT\cite{Serpent2015}, using the nuclear data library ENDF/B-VII.0.  
Serpent can model packed pebbles and TRISO particles explicitly as the way they are manufactured with the locations given by Discrete Element Method(DEM) computation, although studies\cite{Fratoni2007} have shown that the packing configuration does not affect the results, provided the packing fraction is preserved when pebbles near the wall are cut arbitrarily by the geometry definition.

A MATLAB package is developed to automatically read data from Serpent output files and produce temperature/density dependent functions for neutronics parameters of as many neutron energy groups as deemed necessary. Currently, an eight-group structure(Table \ref{tab:egroup}) is used to capture the resonance features in cross sections of the major isotopes: U-235, U-238, and those in flibe (shown in Figure~\ref{fig:xs_isotope}).
\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{./images/diffusion/isotope_XS.png}
  \caption{Cross sections of major isotopes in the core.}
  \label{fig:xs_isotope}
\end{figure}

As shown in Equation \ref{eq:xs_fuel}, temperature dependent cross-sections of the fuel are modeled as log-linear function because the temperature feedback from the fuel is essentially due to Doppler broadening that has a logarithmic dependence on temperature. The temperature/density dependent cross-sections of the flibe are modeled as linear function of the liquid density because the temperature reactivity feedback in the flibe coolant is mainly due to thermal expansion of the liquid and the linear relationship between the density and the macroscopic cross sections.
% The data are calculated at five different fuel temperatures (300K, 600K, 900K, 1200K, 1500K) and five different flibe densities (17, 18, 19, 20, and 21 kg/m$^3$).

\begin{equation}
  \Sigma(T_{fuel}) = c_0 + c_1log(T_{fuel}/T_0)
\label{eq:xs_fuel}   
\end{equation}

\begin{equation}
  \Sigma(T_{flibe}) = c_0 + c_1(\rho_{flibe} - \rho_0)
\label{eq:xs_flibe}   
\end{equation}

$c_0$ and $c_1$ are constant coefficients that can be determined by regression analysis.

\subsection{Thermal-hydraulics}
Full core direct CFD modeling of the flow passage formed by more than 10000 pebbles is not practical with current computation capacity. Porous media model requires considerably less computational resource and is used in the multi-physics model.

An important parameter for porous media is the porosity, $\epsilon$, defined as the ratio between the volume of the void to the total volume(Equation \ref{eq:pf}). It appears in the mass(equation \ref{eq:mass}), momentum(equation \ref{eq:mmtm}) and energy conservation equations(equations \ref{eq:pm_liq} and \ref{eq:pm_sol}) below. Local variation in porosity in a pebble bed, including the drop near the wall due to ordered packing, is not modeled explicitly in the porous media model, but the empirical correlations for pressure loss and heat transfer coefficient incorporate corrections for wall effect.

@@

The momentum equation for fluid flow in the pebble bed is written as:
% comsol
\begin{equation}
  \frac{\rho_f}{\epsilon} \left[(u\nabla)\frac{u}{\epsilon} \right] + \rho\nabla u = \nabla \left[ -pI + (\frac{\mu}{\epsilon}(\nabla u - (\nabla u)^T) - \frac{2\mu}{3\epsilon} \nabla u I)\right] - (\frac{\mu}{k} + \beta_F \|u\| + \rho\nabla u) u + F
  \label{eq:mmtm}
\end{equation}

The dependent variable here is the superficial velocity, the velocity that the fluid would flow if the channel is free of the solid medium. It is defined as the ratio between the volumetric flow rate and the effective cross section area of the flow channel(Equation \ref{eq:sv}).
\begin{equation}
  u=\frac{Q}{A}
  \label{eq:sv}
\end{equation}

@@

The Forcheimer drag coefficient $\beta_F$ and the permeability K are defined in the COMSOL momentum equation according to pebble bed pressure drop characteristics. 
Various pebble bed pressure drop correlations were investigated in previous studies\cite{Scarlat2012} and compared with experimental data\cite{Kang2010}. The most widely used is the Ergun correlation\cite{Ergun1949} for its wide range of validity in laminar, turbulent and transitional regions, its interpretability and convenient implementation, as well as the satisfactory results it provides. It models the pressure loss in a pebble bed as a weighted sum of the viscous energy loss and the inertial energy loss:

@@

They are defined as in equation \ref{eq:betaf} and \ref{eq:K} so that the pressure drop equation in COMSOL(equation \ref{eq:comsol_dp}) matches the Ergun correlation(equation \ref{eq:ergun}).

@@

In a porous media model, the convective heat transfer between the fuel pebbles and the coolant is computed from empirical correlations, whose validation affects directly the credibility of the results. The Wakao correlation(equation \ref{eq:wakao})\cite{Wakao1979} calculates the convective heat transfer coefficient between pebbles and coolant, with Pr and Re the Prandtl and Reynolds numbers.
The constants in the correlation are obtained from a collection of heat transfer data that are carefully examined and corrected for the axial fluid thermal dispersion coefficients. Even though the experimental data that the correlation used to compute the coefficients do not expand the entire FHR flow conditions, recent experimental measurements in randomly packed pebble bed\cite{Huddar2016} have concluded that the Wakao correlation predicts the heat transfer in the FHR conditions adequately.

@@

@@

The geometry is represented in COMSOL as a CAD model, on which a mesh is generated by the COMSOL meshing tools.
Mesh quality is crucial for finite element method(FEM) simulation as the FEM approximates the real solution by solving a descretized equation system over a mesh grid. For most geometry, a free tetrahedral mesh suffice. To further improve the mesh quality with limited element number, hexahedrons can be used in cylindrical components to minimize the number of mesh elements. The mesh size is directly linked to computation cost and result accuracy. As the mesh element size goes to zero, the solution converges to the real solution at the cost of almost infinite computational resources. 
To find a mesh with optimal finesse, a refinement study compares meshes with different sizes, and choose the one with the least computational requirement that satisfies the convergence criteria.

In addition to parameters such as material properties, the code takes temperature dependent cross-section functions as input.

COMSOL provides built-in modules for various physics as well as the possibility for user defined PDEs. Neutron diffusion and $SP_N$ equations are defined in COMSOL through the 'user defined PDEs' interface. The heat transfer and momentum equations are generated directly by COMSOL's porous media and heat transfer models, which were discussed in the previous section.

%(https://www.sharcnet.ca/Software/Ansys/17.0/en-us/help/ans_bas/Hlp_G_BAS3_4.html)

Once the equation system is assembled, the solvers start by searching for the eigenvalue in neutronics equations(Equation \ref{eq:diffusion}) with cross-sections computed from an initial guess of uniform temperatures. 
With the flux distribution found by the eigenvalue solver, a stationary solver computes the temperature distribution in the core by solving the conservation equations in thermal-hydraulics modules. However, an eigenvalue solver evaluate the fluxes and power density values(equation \ref{eq:powerdensity}) on an arbitrary scale, so the power density is normalized onto the operation power in steady state solver , as shown in Equation \ref{eq:scale}.

@@

At the end of the iteration, the fluxes are scaled so that the volume integration of the power density equals to the operating power of the reactor in simulation. New delayed neutron precursor concentrations are computed from the normalized fluxes. The scaling factors are shown in the following equations:
\begin{equation}
    \flux_{gN} = \flux_g \frac{P_{op}}{\iiint_V{\dot{P}dv}}
\end{equation}

@@

where the quantities are defined as follows:\\
$P_{op}$ = operation power, W\\
$\dot{p}$ = volumetric power density, $W/m^3$\\
$\nu$ = mean number of neutrons generated per fission\\
$\Sigma_{f,g}$ = macroscopic fission cross-section of group $g$, m$^{-1}$\\
  $\flux_{g}$ = neutron flux in the $g$-$th$ energy interval [$E_g$, $E_{g-1}$], m$^{-2}$s$^{-1}$
\\
  $\beta_d$ = delayed neutron fraction for delayed neutron precursor group $d$
\\
  $\lambda_d$ = average decay constant of delayed neutron precursors in group d, s$^{-1}$
\\
  $C_d$ = delayed neutron precursors concentration in group d
 \\
 G = total number of energy groups\\
 D = total number of delayed neutron precursors groups.

Once the flux and neutron precursor concentration are scaled, one can simulate transient behaviour by solving the coupled equation system with time derivative terms using COMSOL's transient solver. Time step size is automatically chosen based on the time derivative of the dependent variables.

FEM solvers discretize the space according to the mesh and construct a linear equation system of the basis functions. 
Solving the system of linear equations is a crucial part and the most computationally demanding part of a simulation, for which COMSOL provides two category of methods: direct solvers( MUMPS, PARDISO, and SPOOLES ) and iterative solvers(e.g.  MUMPS, GMRES, FGMRES).
In principal, direct solvers are easier to set up as they do not need preconditioning but require substantial amount of memory for matrix factorization. Iterative solvers approach the solution gradually and requires significantly less memory, but need problem specific tuning.
For smaller mesh, direct solver is faster because it does not need preconditioning and takes a single stride to the solution. But when the memory requirement approaches the hardware limit, the iterative solver performs better by approaching gradually to the solution.
The choice of solvers and their parameters should depend on the specific applications, but all the solvers should find the same results if enough iteration is given.

By default, the model is considered converged when the estimated error in the iterative solver is below $10^{-3}$. Whether this is sufficient for our application depends on the requirement on precision. This error also need to be put in perspective of other uncertainties involved in the modeling process, the default tolerance is enough if it is negligible compare to the errors introduced by the input parameters.

\section{Case study}
\label{sec:res}
The above methodolgoy is applied to the 10~MW Thorium Molten Salt Reactor (TMSR) SF-1 pebble-bed reactor\cite{wang2014a} developed at the Shanghai Institute of Applied Physics (SINAP).

@@

During normal operation, the coolant salt enters at the bottom inlet at 0.18 m/s and 672$^{\circ}C$ and flows upward. The thermo-physical properties of the salt have not been as extensively measured as water. This model uses recommanded values from Romatoski's review of Fluoride salt properties\cite{Romatoski2017a}, which are listed in Table~\ref{tab:flibe}. The impact of uncertainties in these parameters on the figures of merit will be evaluated in the future.
Fuel pebbles are slightly buoyant in the salt and therefore fills the upper region, whereas the lower region of the core is left with flibe salt. 11558 fresh pebbles are loaded in the core in SERPENT simulation with DEM until the core is critical. An outer graphite reflector surrounds the fuel and coolant, providing neutron moderation and reflection. Independent coolant channels are implanted in the reflector to maintain its temperature. Thus the reflector does not participate in the temperature feedback during most transients except for those that involves issues in the reflector coolant channels.

\begin{table}
  \caption{Flibe salt thermo-physical properties.}
  \begin{tabular}[h]{lc}
    \hline
    Property & Value\\
    \hline
    Viscosity, kg/m.s & 1.16E-4*e$^{3755/T[K]}$\\
    Heat capacity, J/kg/K & 2386\\
    Thermal conductivity, W/m/K & 1.1\\
    Density, kg/m$^3$ & 2413 - 0.488T[K]\\
    \hline
  \end{tabular}
\label{tab:flibe}
\end{table}

@@

Coupled heat transfer and neutron diffusion equations are solved with homogenized material properties for each region. Coolant and fuel thermophysical properties are presented in table \ref{tab:flibe}, table \ref{tab:TRISO_prop} and table \ref{tab:pb_prop}. The group constants and delayed neutron parameters are generated from a Serpent model for the TMSR SF-1 core (figure \ref{fig:serpent}) with explicit representation of randomly packed TRISO particles and fuel pebbles. The nuclear data library ENDF/B-VII.0 was used for Monte Carlo neutron transport. 
An eight-group energy structure was chosen with the boundaries shown in table \ref{tab:egroup} to capture the cross section features of the major isotopes: U-235, U-238, and those in flibe(shown in figure \ref{fig:xs_isotope}).

\begin{table}
  \caption{Dimensions and material properties of each layer of TRISO particles.}
  \begin{tabular}[h]{lccccc}
    \hline
    Region&Material&Thickness/Radius,&Density,&Conductivity,&Specific heat,\\
    &&$10^{-6}$ m&kg/m$^3$&W/m/K&J/kg/K\\
    \hline
    Kernel&19.9\% UC$_{0.5}$O$_{1.5}$&250&10,500&3.5&400\\
    Buffer&porous carbon&95&1,000&0.5&2,000\\
    IPyC&pyrolytic carbon&40&1,900&4.0&2,000\\
    SiC&silicon carbide&35&3,200&30&1,300\\
    OPyC&pyrolytic carbon&40&1,900&4.0&2,000\\
    \hline
  \end{tabular}
  \label{tab:TRISO_prop}
\end{table}

>>>>

\begin{abstract}
Fluoride-salt-cooled High-temperature Reactor(FHR) is one of the generation IV reactor designs that promises great safety and market competitiveness because of its unique combination of TRISO fuel elements, liquid salt coolant and graphite reflectors. This paper describes methodologies for developing fully coupled neutronics and thermal hydraulics models for FHRs to incorporate its novel design features, such as the double heterogeneity in the fuel elements.

 To demonstrate its ability to accurately predict the dynamic behavior of a FHR core during postulated transients, which is important for FHR technology development and licensing, an example model is built for the Thorium Melten Salt Reactor Solid Fuel - 1 (TMSR SF-1) design and results from eigenvalue, steady state and transient studies are extensively verify against a Monte Carlo neutron tranport and CFD based reference code.

\end{abstract}

@@

As illustrated in figure \ref{fig:triso}, a commercial scale reactor core would contain tens of thousands fuel pebbles and within each thousands of TRISO particles that enclose the fuel material. This peculiar geometry poses unique challenges to both neutronics and thermal-hydraulics modeling.

Reactor-agnostic Monte Carlo methods can model the particles and pebbles explicitly as the way they are manufactured with the locations given by Discrete Element Method(DEM) computation, although studies\cite{Fratoni2007} have shown that the packing configuration does not affect the results, provided the packing fraction is preserved when pebbles near the wall are cut arbitrarily by the geometry definition. But this approach require a large number of source neutrons to achieve desired statistical accuracy, which implies expensive computational cost, especially for transient analysis. Monte Carlo codes are used in practice to generated cross-sections for deterministic codes and to provide reference results for code verification. Deterministic codes, in the contrary, model the fuel region as homogenized medium with equivalent cross-sections. In particular, diffusion equations can compute two-dimensional or three-dimensional neutron behaviour with reasonable computational cost and accuracy, even though limitations can be found in vicinity of control rods, burnable poisons, and other abrupt discontinuities.

@@

The question of how to model the complex geometry of the flow passage in a pebble bed has been tackled in chemistry engineering \cite{Dixon2001}\cite{Miroliaei2011} for fixed bed reactors in chemical processes and in nuclear engineering for pebble bed reactor designs.  One-dimensional heat conduction simplification is often used in Light Water Reactor (LWR) fuel rods and can be used in spherical fuel elements for FHRs for quick scope analysis, but is inadequate for detailed investigation because it ignores temperature gradients in the axial and the radial direction, which can be significant in a FHR core with cross flow. Two main approaches for full core CFD modeling are realistic modeling and porous media modeling. Realistic modeling has been used to model flow passage directly in a segment of pebbles, from several to several hundreds \cite{Lee2007}\cite{Hassana}\cite{Shams2013}, but quickly becomes unfeasible for a large number of pebbles with the current computation resources. In addition, although the goal of realistic modeling is to model the flow passage realistically, it is not free of assumptions \cite{Lee2007}. For example pebble packing structure and contact point treatment are often used to reduce computational cost. The porous media approach, on the other hand, computes only the average flow conditions with ten time less mesh grids. Comparisons between the direct CFD and porous media methods\cite{Wu2010} suggest that porous media methods provide similar averaged pressure and temperature gradient across the pebble bed to the realistic approach, given the appropriate Darcy-Forchheimer drag model is used. Porous media approach has been used in standard HTR modeling tools Thermix \cite{Struth1985} and Tinte \cite{Gerwin1987}, as well as in a thermal-hydraulics model for FHR \cite{Scarlat2012}.
Appropriate safety margin should be applied in the design given the fact that porous media models do not capture local vertices, flow separation, variations in flow induced heat transfer around the pebble surface and the resulting local hot spots.

Previously, researchers have developed separate codes in thermal-hydraulics \cite{Scarlat2012} and neutronics \cite{Cisneros2013} for FHRs. These codes can be coupled for multiphysics computation using the operator-split method: solving one equation system at a time and transfer data between the codes after each iteration until the results converge. However, the operator split method requires small time steps to ensure convergence due to nonlinear inconsistency \cite{Ragusa2009} and could require a substantial amount of computational resource. This work, instead, explores methodology for developing fully coupled neutronics and thermal-hydraulics models for pebble bed FHRs.

The rest of the paper is organized in three sections. The methodology, including governing equations, generation of cross-sections and implementation in COMSOL is discussed in section \ref{sec:methodology}. A case study based on TMSR design is then presented in section \ref{sec:res}, where model verification procedures via code-to-code comparison is presented and both steady state and transient results are discussed. Lastly, future research needs are identified and discussed in the conclusion.

@@

\subsection{Neutronics}
\subsubsection{Multi-group neutron diffusion equations}
Multi-group diffusion equations, which provide a good balance between CPU cost and accuracy, is used in the multiphyscis model for neutronics computation. The governing equations of multi-group neutron diffusion with delayed neutrons are written as following:

\begin{align}
    \frac{1}{v_g}\frac{\partial \phi_g}{\partial t} = \nabla D_g \nabla \phi_g - \Sigma_{t,g}\phi_g + \sum_{g'=1}^{G} \Sigma_{s, g'g}\phi_{g'} +  \nonumber \\
    (1-\beta)\chi_{p,g}\sum_{g'=1}^G(\nu \Sigma_f)_{g'} \phi_{g'} + \chi_{d,g}\sum_{i=1}^D \lambda_iC_i
\label{eq:diffusion}
  \end{align}

@@

For criticality calculations, the effective multiplication factor \keff is inserted into the neutron balance equation to represent the variation of neutron population from one generation to the following one.

% , resulting in:
%   \begin{align}
%     0 =& \nabla D_g \nabla \phi_g - \Sigma_{a,g}\phi_g - \sum_{\substack{1\\g'\neq g}}^G \Sigma_{s, gg'}\phi_{g} + \\\nonumber &\sum_{\substack{1\\g'\neq g}} \Sigma_{s, g'g}\phi_{g'} + \frac{\chi_{t,g}}{k_{\mathrm{eff}}}\sum_{g'=1}^G(\nu \Sigma_f)_{g'} \phi_{g'}
%   \end{align}

\subsubsection{Cross-section generation}
The diffusion model uses group-averaged, component-wise neutronic constants, generated with the Monte Carlo neutron transport code.
In order to perform transient analysis, these cross-sections need to be paramatrized as functions of reactor operating conditions, notably fuel temperature and coolant density in a FHR core, via a process called computer experiments, in which we run a Monte Carlo model with a sample of operating conditions and fits a function between the resulting cross-sections and the input data sets.

The change in flibe cross-sections with temperature is mainly due to the change in density of liquid flibe, thus the temperature dependent function for flibe cross-sections are in the following form:

\begin{align}
  \hat{\Sigma}(\rho_{flibe}) = c_0 + c_1(\rho_{flibe} - \rho_0)
  \label{eq:flibe_xs}
\end{align}

where $\hat{\Sigma}(\rho_{flibe})$ is the estimation of a flibe macroscopic cross-section, $\rho_{flibe}$ is the flibe density, $\rho_0$ is the nominal flibe density, and 
$c_0$ and $c_1$ are the constants to be calculated.

Unlike the liquid salt coolant, the solid fuel do not change shape or density in the range of the considered temperatures. The fuel temperature reactivity feedback (also called the Doppler feedback) is mainly due to the Doppler broadening of the microscopic cross-sections of the fuel material.
Therefore, the fuel cross-sections are described by a linear-log function of the fuel temperature. During fast transients such as reactivity induced accidents, the heat does not have time to reach the coolant. The core relies on the temperature reactivity feedback from the fuel inside the TRISO particles to stabilize. It is important to compute the temperature distribution inside TRISO particles and fuel pebbles, and use the temperatures at different depth to compute cross-sections. The cross-sections in the fuel region is expressed in the following form:

\begin{align}
  \hat{\Sigma} (T_{fuel}) = c_0 + c_1(log(T_{fuel}) - log(T_0))
  \label{eq:XS_linear_log}
\end{align}

where $\hat{\Sigma}(T_{fuel})$ is the estimation of a fuel macroscopic cross-section, $T_{fuel}$ is the fuel temperature, or temperature vector in case many temperatures affect the homogenized cross-section in a fuel region, and $c_0$, $c_1$ are the coefficients we search for.

\subsection{Thermal-hydraulics}
Direct CFD modeling of the flow passage formed by more than 10000 pebbles is not practical with current computation capacity. Porous media model requires considerably less computational resource and is used in the multi-physics model.

An important parameter for porous media is the porosity, $\epsilon$, defined as the ratio between the volume of the void to the total volume (Equation \ref{eq:pf}). It appears in the mass (Equation \ref{eq:mass}), momentum (Equation \ref{eq:mmtm}) and energy conservation equations (equations \ref{eq:pm_liq} and \ref{eq:pm_sol}) below. Local variation in porosity in a pebble bed, including the drop near the wall due to ordered packing, is not modeled explicitly in the porous media model, but the empirical correlations for pressure loss and heat transfer coefficient incorporate corrections for wall effect.

@@

The momentum equation for fluid flow in the pebble bed is written as:
% comsol
\begin{align}
  \frac{\rho_f}{\epsilon} \left[(u\nabla)\frac{u}{\epsilon} \right] + \rho\nabla u = &\nabla \left[ -pI + (\frac{\mu}{\epsilon}(\nabla u - (\nabla u)^T) - \frac{2\mu}{3\epsilon} \nabla u I)\right]  \nonumber\\
  &- (\frac{\mu}{k} + \beta_F \|u\| + \rho\nabla u) u + F
  \label{eq:mmtm}
\end{align}

The dependent variable here is the superficial velocity, the velocity that the fluid would flow if the channel is free of the solid medium. It is defined as the ratio between the volumetric flow rate and the effective cross section area of the flow channel (Equation \ref{eq:sv}).
\begin{equation}
  u=\frac{Q}{A}
  \label{eq:sv}
\end{equation}

@@

The Forcheimer drag coefficient $\beta_F$ and the permeability K are defined in the COMSOL momentum equation according to pebble bed pressure drop characteristics. 
Various pebble bed pressure drop correlations were investigated in previous studies \cite{Scarlat2012} and compared with experimental data \cite{Kang2010}. The most widely used is the Ergun correlation \cite{Ergun1949} for its wide range of validity in laminar, turbulent and transitional regions, its interpretability and convenient implementation, as well as the satisfactory results it provides. It models the pressure loss in a pebble bed as a weighted sum of the viscous energy loss and the inertial energy loss:

@@

They are defined as in equation \ref{eq:betaf} and \ref{eq:K} so that the pressure drop equation in COMSOL (equation \ref{eq:comsol_dp}) matches the Ergun correlation (equation \ref{eq:ergun}).

@@

In a porous media model, the convective heat transfer between the fuel pebbles and the coolant is computed from empirical correlations, whose validation affects directly the credibility of the results. The Wakao correlation (equation \ref{eq:wakao}) \cite{Wakao1979} calculates the convective heat transfer coefficient between pebbles and coolant, with Pr and Re the Prandtl and Reynolds numbers.
The constants in the correlation are obtained from a collection of heat transfer data that are carefully examined and corrected for the axial fluid thermal dispersion coefficients. Even though the experimental data that the correlation used to compute the coefficients do not expand the entire FHR flow conditions, recent experimental measurements in randomly packed pebble bed \cite{Huddar2016} have concluded that the Wakao correlation predicts the heat transfer in the FHR conditions adequately.

@@

\subsubsection{Multi-scale temperature distribution}

In order to determine the temperature distribution inside fuel pebbles and TRISO particles, which are crucial for accurately determining temperature feedback effect from the fuel material during fast transient analysis, we use the 1-D heat diffusion equation in spherical coordinate to model the heat conduction inside solid fuel elements:

\begin{align}
  \rho C_p\frac{\partial T}{\partial t} = \frac{1}{r^2}\frac{\partial}{\partial r}\left(kr^2\frac{\partial T}{\partial r}\right) + g
\end{align}
where r is the radial coordinate of the point, $\rho$ is density, $C_p$ is specific heat capacity, T is temperature, t is time, k is thermal conductivity, and g is the heat generation density.

And the resulting equation from a change of variable $U(r,t)=rT(r,t)$ is shown below:

\begin{align}
  \frac{\partial U}{\partial t} = \alpha \frac{\partial ^2 U}{\partial r^2} + r \frac{g}{\rho C_p}
\end{align}

with boundary conditions:
\begin{align}
U_0 = 0\\
\frac{\partial U}{\partial r} |_{r=R} = \left( \frac{1}{R} - \frac{h}{k}\right) U + \frac{R}{k}hT_{\infty}
\end{align}

This equation system is readily solved numerically under finite volume discretization, where spherical elements are divided into fine annuli, so that the temperature in each can be considered uniform.

Likewise, the coated particle is divided into n layers. In order to save computational cost, the coatings are combined into a non-heat-generating layer with equivalent heat transfer properties. The heat transfer equation for the jth layer in a TRISO particle that is situated in the ith layer of a fuel pebble is shown in equation \ref{eq:ij_diffusion}.

\begin{equation}
    \frac{\partial{Tp_{i,j}}}{\partial{t}}\rho_{i,j}c_{p_{i,j}} = \frac{k_{i,j}}{r_{i,j}}   \Bigg[  \frac {\frac{r_{i,j+1}T_{i,j+1}}{r_{i,j+1}-r_{i,j}}+\frac{r_{i,j-1}T_{i, j-1}}{r_{i,j}-r_ {i, j-1}  }}{\frac{1}{2}(r_ {i, j+1} -r_ {i, j-1} )}-\frac{2r_ {i, j} T_ {i,j} }{(r_ {i,j} -r_ {i,j-1} )(r_ {i, j+1} -r_ {i,j})} \Bigg]  + g  
    \label{eq:ij_diffusion}
\end{equation}

where r is the radius of TRISO particles.

@@

Once the equation system is assembled, the solvers start by searching for the eigenvalue in neutronics equations (equation \ref{eq:diffusion}) with cross-sections computed from an initial guess of uniform temperatures. 
With the flux distribution found by the eigenvalue solver, a stationary solver computes the temperature distribution in the core by solving the conservation equations in thermal-hydraulics modules. However, an eigenvalue solver evaluate the fluxes and power density values (equation \ref{eq:powerdensity}) on an arbitrary scale, so the power density is normalized onto the operation power in steady state solver , as shown in equation \ref{eq:scale}.

@@

At the end of the iteration, the fluxes are scaled so that the volume integration of the power density equals to the operating power of the reactor in simulation. Normalized delayed neutron precursor concentrations are also computed, from the normalized fluxes. The scaling factors for both are shown in the following equations:
\begin{equation}
    \flux_{gN} = \flux_g \frac{P_{op}}{\iiint_V{\dot{P}dv}}
\end{equation}

@@

where the quantities are defined as follows:\\
$P_{op}$ = operation power, W\\
$\dot{P}$ = volumetric power density, $W/m^3$\\
$\nu$ = mean number of neutrons generated per fission\\
$\Sigma_{f,g}$ = macroscopic fission cross-section of group $g$, m$^{-1}$\\
  $\flux_{g}$ = neutron flux in the $g$-$th$ energy interval [$E_g$, $E_{g-1}$], m$^{-2}$s$^{-1}$
\\
  $\beta_d$ = delayed neutron fraction for delayed neutron precursor group $d$
\\
  $\lambda_d$ = average decay constant of delayed neutron precursors in group d, s$^{-1}$
\\
  $C_d$ = delayed neutron precursors concentration in group d
 \\
 G = total number of energy groups\\
 D = total number of delayed neutron precursors groups.

Once the flux and neutron precursor concentration are scaled, one can simulate transient behaviour by solving the coupled equation system with time derivative terms in COMSOL's transient solver, with time step size being automatically chosen based on the time derivative of the dependent variables.

\section{Case study}
\label{sec:res}
The above methodolgoy is applied to a 10~MW research reactor design, the Thorium Molten Salt Reactor (TMSR) SF-1 \cite{wang2014a}, described in the following sections.

\subsection{TMSR SF-1 design overview}

@@

During normal operation, the coolant salt enters at the bottom inlet at 0.18 m/s and 672$^{\circ}C$ and flows upward. 
Fuel pebbles are slightly buoyant in the salt and therefore fills the upper region, whereas the lower region of the core is left with flibe salt. 11558 fresh pebbles are loaded in the core in SERPENT simulation with DEM until the core is critical. An outer graphite reflector surrounds the fuel and coolant, providing neutron moderation and reflection. Independent coolant channels are implanted in the reflector to maintain its temperature. Thus the reflector does not participate in the temperature feedback during most transients except for those that involves issues in the reflector coolant channels.

@@

\subsection{Multiphysics modeling for TMSR SF-1}

As shown in Figure \ref{fig:model_geom}, the multi-physics reactor core model is composed by three regions: an upper porous region that represents a fuel pebble and flibe salt porous mixture, a lower region with only flibe salt and an outer graphite reflector region. Coupled heat transfer and neutron diffusion equations are solved with homogenized material properties for each region.

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.2\textwidth}
    \includegraphics[width = \columnwidth]{./images/diffusion/tmsr/2d_schematic.png}
    %\caption{Schematic of the TMSR core model geometry}
  \end{subfigure}
  ~
  \begin{subfigure}{0.55\textwidth}
    \includegraphics[width = \columnwidth]{./images/diffusion/tmsr/finer.png}
    %\caption{Schematic of the 2D mesh}
  \end{subfigure}
  \caption{Schematic of the multi-physics model for TMSR SF-1 and a three-dimensional mesh}
  \label{fig:model_geom}
\end{figure}

Multigroup neutron diffusion equations are implemented into COMSOL as user defined PDEs, with the homogenized group constants for each regions defined as input. Furthermore, $SP_N$ treatment can be used in regions where the diffusion assumption is questioned, such as in the vicinity of the control rods.
The group constants that are used by the diffusion and $SP_N$ models are generated from a Serpent model for the TMSR SF-1 core with explicit representation of randomly packed TRISO particles and fuel pebbles, using the nuclear data library ENDF/B-VII.0.
